---
title: 'Week 03 - Assignment: Electricity Market Prices'
author: "Uğur Aşkar"
date: "2020-11-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(knitr)
```

## Preparing data for analysis

```{r}
raw_df = read.csv("C:\\data\\EPIAS_EXIST_MCP_SMP\\ptf-smf.csv",header=TRUE,dec=",")
knitr::kable(raw_df[1:5,], format="markdown", align='cc')
```
Pulling data from our file system with read_csv function.

```{r, warning=FALSE}
char_to_numeric = function(char_string){
  numeric_value = as.numeric(sub(',', '.',sub('.', '', char_string, fixed = TRUE)))
  return(numeric_value)
}

clean_names = c("DATE_TIME","MARKET_CLEARING_PRICE","SYSTEM_MARGINAL_PRICE","POSITIVE_IMBALANCE_PRICE","NEGATIVE_IMBALANCE_PRICE","SYSTEM_MARGINAL_PRICE_COURSE")
clean_df = setNames(raw_df, clean_names) %>% transform(DATE_TIME = as.POSIXct(DATE_TIME,format='%d.%m.%y %H:%M'), SYSTEM_MARGINAL_PRICE = char_to_numeric(SYSTEM_MARGINAL_PRICE), NEGATIVE_IMBALANCE_PRICE = char_to_numeric(NEGATIVE_IMBALANCE_PRICE))
knitr::kable(clean_df[1:5,], format="markdown", align='cc')
```
Cleaned our data here also defined a function to convert our data types to numeric(instead of char) so we can analyze it further.

```{r}
working_time = function(week_day, hour){
    isItWorkingTime = ifelse(between(week_day,2,6) & between(hour,8,18), "Working Time", "Free Time")
    return(isItWorkingTime)
  }


analytical_df = clean_df[1:3] %>% transmute(DATE_TIME, HOUR = hour(DATE_TIME), WEEK_DAY = wday(DATE_TIME, label = TRUE, abbr=FALSE, locale='UK'), WORKING_TIME = working_time(wday(DATE_TIME),hour(DATE_TIME)), MARKET_CLEARING_PRICE, SYSTEM_MARGINAL_PRICE)
knitr::kable(analytical_df[1:5,], format="markdown", align='cc')
```

## Visualizing the data
```{r}
ggplot(analytical_df, aes(x=DATE_TIME, group=1)) +
  geom_line(aes(y = MARKET_CLEARING_PRICE, color = "Market Clearing Price") , size = 1) +
  geom_line(aes(y = SYSTEM_MARGINAL_PRICE, color = "System Marginal Price"), size = 1) +
  theme_minimal() + 
  labs(x = '', y = 'Price in TRY/MWh', color = "Acquired By", title = "MCP/SMP", subtitle = "Electricity market prices in September 2020")
```

Here we can see that there are spiking in System Marginal Prices in 03/09/2020, 07/09/2020 and 17/09/2020 but no national anomaly detected in news search. There must be a hidden variable we can't see.

```{r}
analytical_df %>%
  select(WORKING_TIME, MARKET_CLEARING_PRICE, SYSTEM_MARGINAL_PRICE) %>%
  group_by(WORKING_TIME) %>%
  summarize(MARKET_CLEARING_PRICE = mean(MARKET_CLEARING_PRICE), SYSTEM_MARGINAL_PRICE = mean(SYSTEM_MARGINAL_PRICE)) %>% ggplot(., aes(x="", y=SYSTEM_MARGINAL_PRICE, fill=WORKING_TIME)) +
  geom_bar(stat="identity") +
  coord_polar("y") + 
  theme_minimal()
```

Here we can see there is more electric consumption in work hours.

## Sources