---
title: "Spam Data"
author: "Uğur Aşkar"
date: "2020-12-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(rpart)
library(rpart.plot)
```

## Preparing data for analysis

```{r, warning = FALSE, error = FALSE, message=FALSE}
raw_df = read_csv("data/spambase.data", col_names = FALSE)
clean_names = c("FREQ_MAKE","FREQ_ADDRESS","FREQ_ALL","FREQ_3D","FREQ_OUR","FREQ_OVER","FREQ_REMOVE","FREQ_INTERNET","FREQ_ORDER","FREQ_MAIL","FREQ_RECEIVE","FREQ_WILL","FREQ_PEOPLE","FREQ_REPORT","FREQ_ADDRESSES","FREQ_FREE","FREQ_BUSINESS","FREQ_EMAIL","FREQ_YOU","FREQ_CREDIT","FREQ_YOUR","FREQ_FONT","FREQ_000","FREQ_MONEY","FREQ_HP","FREQ_HPL","FREQ_GEORGE","FREQ_650","FREQ_LAB","FREQ_LABS","FREQ_TELNET","FREQ_857","FREQ_DATA","FREQ_415","FREQ_85","FREQ_TECHNOLOGY","FREQ_1999","FREQ_PARTS","FREQ_PM","FREQ_DIRECT","FREQ_CS","FREQ_MEETING","FREQ_ORIGINAL","FREQ_PROJECT","FREQ_RE","FREQ_EDU","FREQ_TABLE","FREQ_CONFERENCE","FREQ_;","FREQ_(","FREQ_[","FREQ_!","FREQ_$","FREQ_#","LENGTH_AVERAGE","LENGTH_LONGEST","LENGTH_TOTAL","IS_SPAM")
clean_df = setNames(raw_df, clean_names)
knitr::kable(head(clean_df[c(1,2,3,4,5,54,55,56,57,58)], 5), format="markdown", align='cc')
```

## Splitting dataframe for train and testing

We are going to split the dataframe by 85% to train dataset and use remaining 15% to control and test our prediction rate. 

```{r}
sample_size = floor(0.85 * nrow(clean_df))
set.seed(9999)

train_index <- sample(seq_len(nrow(clean_df)), size = sample_size)

train_df = clean_df[train_index, ]
test_df = clean_df[-train_index, ]
knitr::kable(head(test_df[c(1,2,3,4,5,54,55,56,57,58)], 5), format="markdown", align='cc')
```
## Building Classification Tree

```{r}
classification_tree = rpart(IS_SPAM~., data = train_df, method='class')
rpart.plot(classification_tree, type=3, box.palette = "Blues")
```
## CP Table of Classification Tree

We will check CP Table of classification tree before we prune the tree. It is not necessary to prune it here since our dataset is small but It is good practice to prune it for performance optimization.

```{r}
cls_tree_cpt = classification_tree$cptable
knitr::kable(cls_tree_cpt, format="markdown", align='cc')
```
## Measuring our prediction rate

```{r}
predict_df = predict(classification_tree, test_df, type = 'class')
prediction_table = table(test_df$IS_SPAM, predict_df)

prediction_rate = (prediction_table[1,1]+prediction_table[2,2])/sum(prediction_table)
print(paste(c('Our CART model has ',round(prediction_rate*100,2),'% correct prediction rate.'), collapse = ""))
```

## References

[How to split data into training/testing sets using sample function](https://stackoverflow.com/questions/17200114/how-to-split-data-into-training-testing-sets-using-sample-function)

[Regression Tree](https://www.rdocumentation.org/packages/rpart/versions/4.1-15/topics/rpart)